---
title: "Lab#1 - Getting Started with R & Rstudio"
author: "Rasmus W. Schmøkel"
date: '2022-09-05'
output: html_document
---

Welcome to the R tutorial of 'Social Data Science'! Exercise sessions and self-study tasks will help you to apply the knowledge learned in class and gain experience.

# Introduction to R

R is a free software environment for statistical computing and graphics. It compiles and runs on a wide variety of UNIX platforms, Windows and MacOS. In this course we will be using R in the RStudio environment to perform our exercises. Exercise sheets and solutions will be R Markdown documents that combine and format text, code, and code output.

Follow the instructions from today's readings to download and install R and RStudio on your computer:

Once you have successfully installed R and RStudio, you can follow through the basic commands below to get familiar with R.


**Remember that there is a great community using R and you can always search online for the way to do things - google is your friend**



# Getting help

If you don't know what a function does, try searching for help with "?"
```{r echo=TRUE}
? mean
```

If you don't know how to do something, your fellow programmer is your friend. For example:

https://stats.stackexchange.com/questions/157661/how-to-calculate-mean-median-mode-std-dev-from-distribution

This is a really useful search engine (Google plus filter for R-relevant content): https://rseek.org/

or just google.com

# Simple operations

You can use either "<-" or "=" to assign a value to a variable.

```{r echo=TRUE}
a <- 5
b = c(2,4,6,8)
d <- c(3,5,7,9)
e <- 7
```
The "c" in the above vector assignment stands for combine into a vector. The elements in the array are indexed from 1 to n. (in some other programming languages the index starts from 0)

You can see the result by typing the variable name to the console.
```{r}
a
b
```

Selecting the 3rd element of the vector:
```{r}
d[3]
```

Now we can do some arithmetic with the vectors.

```{r}
a*b
```

```{r}
b+d
b+1
```

Notice when a vector is multiplied with a scalar (=a single value/number), each of the elements in the vector is multiplied. And when two vectors are added/multiplied, they must be the same dimensions and the arithmetic happens elementwise.

Boolean values can also be stored and manipulated in R

```{r}
b1 <- TRUE
b2 <- 1>2
```


rep, seq, and rev are useful functions to produce and manipulate simple vectors

```{r}
rep(1,7)
```

```{r}
seq(1,7,by=2)
```

```{r}
rev(seq(1,7))
```

Your turn:

**1) We create a vector V with the even numbers between 1 and 10. Show its content.**
```{r}
#Your code here

```


**2) look at its third entry and test if it is larger than 3.**
```{r}
#Your code here

```

**3) Reverse its order and divide each of its entries by 2.**
```{r}
#Your code here
```

**4) Check whether the value of a is higher than the value of  object 'e'**
```{r}
#Your code here


```

**5) Check whether 'alpha' and 'beta' is equal**
```{r}


alpha <- round(log(150))
beta <- 50/10

#Your code here


```


**8) Install the package 'ggplot2'**
```{r}

#Your code here
```
**9) Load the package 'ggplot2'**
```{r}

#Your code here
```

**10) ggplot2 is used for plotting basically everything - and done beautifully. We will return to data visualization using ggplot2 in lesson 3. Until then, here is a basic plot using the built-in dataset 'mtcars'**
```{r}
ggplot() + geom_point(data = mtcars, aes(x = disp, y = mpg))
```
